Draft

### 2022-06-20 Final/FinalReport.qmd
Update FinalReport.qmd
david's changes. issues: natl pride in CI and predict; CI dealing with the factor variables; remove recipes and workflows for models 2,3

```{r}
#| label: load-pkgs
#| message: false
#| warning: false
library(tidyverse)
library(tidymodels)
library(knitr)
library(ggfortify)
library(GGally)
```

```{r load-data}
#| message: false
abortion_data_full<-read_csv(here::here("data/abortion-attitudes", 
                                        "wvs-usa-abortion-attitudes-data.csv"))
```

@@

```{r visualization-summary}
#| warning: false
ggplot(abortion_data_full, aes(x = aj)) +
  geom_histogram(binwidth = 1, aes(y=..density..)) + 
  geom_density(color = "red") +
  labs(title = "Histogram of Abortion Attitudes",
       x = "Abortion Justifiability (1-Least, 10-Most)")

summary(abortion_data_full$aj)
```

@@

```{r eda-year}
#| warning: false
#Year of survey
ggplot(abortion_data_full, aes(x = year)) +
  geom_histogram() + 
  labs(title = "Histogram of year")
ggplot(abortion_data_full, aes(group = year, y = aj)) +
  geom_boxplot() 
```

```{r eda-ideology}
#| warning: false
# ideology
ggplot(abortion_data_full, aes(x = ideology)) +
  geom_histogram() + 
  labs(title = "Histogram of ideology")
ggplot(abortion_data_full, aes(x = ideology,group = ideology, y = aj)) +
  geom_boxplot()
```

```{r eda-cai}
#| warning: false
# Child autonomy index
ggplot(abortion_data_full, aes(x = cai)) +
  geom_histogram() + 
  labs(title = "Histogram of Child autonomy index")
ggplot(abortion_data_full, aes(x = cai, group = cai, y = aj)) +
  geom_boxplot()
```

```{r eda-god}
#| warning: false
# Importance of God
ggplot(abortion_data_full, aes(x = godimportant)) +
  geom_histogram() + 
  labs(title = "Histogram of how respondent saw God's importance")
ggplot(abortion_data_full, aes(x = godimportant, group = godimportant, 
                               y = aj)) +
  geom_boxplot()
```

```{r eda-respect}
#| warning: false
# Respect for authority
ggplot(abortion_data_full, aes(x = respectauthority)) +
  geom_histogram() + 
  labs(title = "Histogram of respect for authority")
ggplot(abortion_data_full, aes(x = respectauthority, group = respectauthority, 
                          y = aj)) +
  geom_boxplot()
```

```{r eda-pride}
#| warning: false
# National Pride
ggplot(abortion_data_full, aes(x = nationalpride)) +
  geom_histogram() + 
  labs(title = "Histogram of National pride")
ggplot(abortion_data_full, aes(x = nationalpride, group = nationalpride, y = aj)) +
  geom_boxplot()
```

```{r correlations}
#| warning: false
ggpairs(abortion_data_full, 
  columns = c("year", "ideology", "cai", "godimportant", "respectauthority", 
              "nationalpride"),
  columnLabels = c("year", "ideology", "child autonomy", "imptnce of God", 
                   "resp for autho", "natl pride")) +
   theme(
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(angle = 45, size = 10),
    strip.text.y = element_text(angle = 0, hjust = 0)
    )
```

@@

@@

```{r clean-data}
abortion_data <- abortion_data_full %>% 
  filter(year >= "1995") %>%
  select(-starts_with("wave"), -starts_with("wvsccode"))

abortion_data$year <- as.factor(abortion_data$year)
abortion_data$respectauthority <- as.factor(abortion_data$respectauthority)
abortion_data$respectauthority <- factor(abortion_data$respectauthority, levels=c('0', '-1', '1'))
abortion_data$nationalpride <- as.factor(abortion_data$nationalpride)
```

@@

```{r splits}
set.seed(206)
abortion_split <- initial_split(abortion_data)
abortion_training <- training(abortion_split)
abortion_testing <- testing(abortion_split)
```

```{r spec}
abortion_spec <- linear_reg() %>%
  set_engine("lm")
```

```{r recipe1}
abortion_rec <- recipe(aj ~ year + ideology + cai + godimportant +
                        respectauthority + nationalpride,
                        data = abortion_data) %>%
  step_center(ideology) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())

@@

```{r recipe2}
abortion_rec2 <- recipe(aj ~ year + godimportant,
                        data = abortion_data) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())

@@

```{r recipe3}
abortion_rec3 <- recipe(aj ~ year + ideology + cai + nationalpride,
                        data = abortion_data) %>%
  step_center(ideology) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())

@@

```{r workflows}
abortion_wflow <- workflow() %>%
  add_model(abortion_spec) %>%
  add_recipe(abortion_rec)

@@

```{r fit-models}
abortion_fit <- abortion_wflow %>%
  fit(data = abortion_training)
tidy(abortion_fit) %>%
  kable(digits = 3)

@@

```{r fit-model}
#| echo: false
abortion_fit <- abortion_wflow %>%
  fit(data = abortion_training)
tidy(abortion_fit) %>%
  kable(digits = 3)
```

@@

$$y ~=~ 8.5 ~+~ 0.029x_1 ~+~ 0.469x_2 ~-~ (0.284x_3 ~+~ 0.288x_4 ~+~ 0.19x_5)$$

where $x_1 ~=~ \text{year}$, $x_2 ~=~ \text{child autonomy index}$, $x_3 ~=~ \text{ideology}$, $x_4 ~=~ \text{god is important}$, and $x_5 ~=~ \text{respect for authority}$.

@@

Inference: CI and HT

We will take a quick look at the 95% confidence intervals of the parameters from our fitted model, displayed in the following plots:

[plots of parameters’ .estimate against the parameter names]

```{r inference}
tidy(abortion_fit, conf.int = TRUE) %>%
  kable(digits = 3)
```

We then proceed to our hypothesis test. As stated in our introduction, given that this issue has become highly polarized by political party, we suspect that liberal political ideology and political attitudes will correlate with belief that abortion is more justified.

@@

HA (alternative hypothesis: An individual’s political ideology (ideology), their beliefs about child autonomy (cai), the importance of God in their life (godimportant), their view on whether respect for authority is good (respectauthority), and their pride in being American (nationalpride) are significant predictors of aj, and the variables’ respective slope point estimates will not be 0.

@@

Prediction:

We now consider the strength of our model in predicting the attitude toward the justifiability of abortion of a person of certain characteristics.

For a person who [describe liberal characteristics here], our model predicts that their attitude toward the justifiability of abortion is [write what the model would predict].

This does/does not support our hypothesis based on contextual knowledge of this issue.

However, calculating the R-squared and root-mean-squared-error (RMSE) values of our model shows that our model is not very strong at predicting attitude toward the justifiability of abortion, with a R^2 of __ and RMSE of __:

[R^2, RMSE output]

```{r predict}
abortion_training_pred <- predict(abortion_fit, abortion_training) %>%
  bind_cols(abortion_training %>% select(aj, year, ideology, cai, 
                                         godimportant, respectauthority, 
                                         nationalpride))
abortion_training_pred
```

@@

@@

```{r cv}
set.seed(206)
folds <- vfold_cv(abortion_training, v = 10)
abortion_fit_rs <- abortion_wflow %>%
  fit_resamples(folds)

@@

```{r cv-table}
cv_metrics <- collect_metrics(abortion_fit_rs, summarize = FALSE) 
cv_metrics2 <- collect_metrics(abortion_fit_rs2, summarize = FALSE) 
cv_metrics3 <- collect_metrics(abortion_fit_rs3, summarize = FALSE)

@@

```{r aic-bic}
glance(abortion_fit) %>% 
  select(AIC, BIC)

>>>>

```{r}
#| label: load-pkgs
#| message: false
#| warning: false
#| echo: false
library(tidyverse)
library(tidymodels)
library(knitr)
library(ggfortify)
library(GGally)
library(cowplot)
library(readr)
```

```{r load-data}
#| message: false
#| echo: false
abortion_data_full<-read_csv(here::here("data/abortion-attitudes", 
                                        "wvs-usa-abortion-attitudes-data.csv"))
```

```{r clean-data}
#| echo: false
abortion_data <- abortion_data_full %>% 
  filter(year >= "1995") %>%
  select(-starts_with("wave"), -starts_with("wvsccode"))
```

@@

```{r visualization-summary}
#| warning: false
#| echo: false
ggplot(abortion_data, aes(x = aj)) +
  geom_histogram(binwidth = 1, aes(y=..density..), fill = "light blue") + 
  geom_density(color = "red") +
  labs(title = "Histogram of Abortion Attitudes",
       x = "Abortion Justifiability (1-Least, 10-Most)")
```

@@

```{r eda-histograms}
#| warning: false
#| message: false
#| echo: false
h1 <- ggplot(abortion_data, aes(x = year)) +
  geom_histogram(fill = "orange") + 
  labs(x = "Year of Survey")

h2 <- ggplot(abortion_data, aes(x = ideology)) +
  geom_histogram(fill = "orange") + 
  labs(x = "Political Ideology")

h3 <- ggplot(abortion_data, aes(x = cai)) +
  geom_histogram(fill = "orange") + 
  labs(x = "Child Autonomy Index")

h4 <- ggplot(abortion_data, aes(x = godimportant)) +
  geom_histogram(fill = "orange") + 
  labs(x = "Importance of God")

h5 <- ggplot(abortion_data, aes(x = respectauthority)) +
  geom_histogram(fill = "orange") + 
  labs(x = "Respect for Authority")

plot_grid(h1, h2, h3, h4, h5, ncol = 3, nrow = 2)

```

```{r eda-boxplots}
#| warning: false
#| echo: false
b1 <- ggplot(abortion_data, aes(x = year, group = year, y = aj)) +
  geom_boxplot() +
  labs(y = "Abortion Attitude", x = "Year of Survey ")

b2 <- ggplot(abortion_data, aes(x = ideology,group = ideology, y = aj)) +
  geom_boxplot()+
  labs(y = "Abortion Attitude", x = "Political Ideology")

b3 <- ggplot(abortion_data, aes(x = cai, group = cai, y = aj)) +
  geom_boxplot()+
  labs(y = "Abortion Attitude", x = "Child Autonomy Index")

b4 <- ggplot(abortion_data, 
             aes(x = godimportant, group = godimportant, y = aj)) +
  geom_boxplot()+
  labs(y = "Abortion Attitude", x = "Importance of God")

b5 <- ggplot(abortion_data, aes(x = respectauthority, group = respectauthority, 
                          y = aj)) +
  geom_boxplot()+
  labs(y = "Abortion Attitude", x = "Respect For Authority")

plot_grid(b1, b2, b3, b4, b5, ncol = 3, nrow = 2)
```

```{r correlations}
#| warning: false
#| echo: false
ggpairs(abortion_data, 
  columns = c("year", "ideology", "cai", "godimportant", "respectauthority"),
  columnLabels = c("YS", "PI", "CAI", "IG", 
                   "RA")) +
   theme(
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(angle = 45, size = 9),
    strip.text.y = element_text(angle = 0, hjust = 0)
    )
```

@@

```{r clean-data2}
#| echo: false
abortion_data$year <- as.factor(abortion_data$year)
abortion_data$respectauthority <- as.factor(abortion_data$respectauthority)
abortion_data$respectauthority <- factor(abortion_data$respectauthority, levels=c('0', '-1', '1'))
abortion_data$nationalpride <- as.factor(abortion_data$nationalpride)
```

@@

@@

```{r splits}
#| echo: false
set.seed(206)
abortion_split <- initial_split(abortion_data)
abortion_training <- training(abortion_split)
abortion_testing <- testing(abortion_split)
```

```{r spec}
#| echo: false
abortion_spec <- linear_reg() %>%
  set_engine("lm")
```

```{r recipe1}
#| echo: false
abortion_rec <- recipe(aj ~ year + ideology + cai + godimportant +
                        respectauthority + nationalpride,
                        data = abortion_data) %>%
  step_center(ideology) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())

@@

```{r recipe2}
#| echo: false
abortion_rec2 <- recipe(aj ~ year + godimportant,
                        data = abortion_data) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())

@@

```{r recipe3}
#| echo: false
abortion_rec3 <- recipe(aj ~ year + ideology + cai + nationalpride,
                        data = abortion_data) %>%
  step_center(ideology) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())

@@

```{r workflows}
#| echo: false
abortion_wflow <- workflow() %>%
  add_model(abortion_spec) %>%
  add_recipe(abortion_rec)

@@

```{r fit-models}
#| echo: false
abortion_fit <- abortion_wflow %>%
  fit(data = abortion_training)
tidy(abortion_fit) %>%
  kable(digits = 3)

@@

```{r fit-model}
#| echo = false
abortion_fit <- abortion_wflow %>%
  fit(data = abortion_training)
tidy(abortion_fit) %>%
  kable(digits = 3)
```

@@

$$\begin{align}y ~=~ &6.596 ~+~ 0.468x_1 ~+~ 0.33x_2 ~+~ 0.339x_3 ~+~ 0.521x_4 ~+~ 0.14x_5 \\&~-~ (0.28x_6 ~+~ 0.287x_7 ~+~ 0.179x_8 ~+~ 0.11x_9)\end{align}$$

where $x_1 ~=~ \text{child autonomy index}$, $x_2 ~=~ \text{year 1999}$, $x_3 ~=~ \text{year 2006}$, $x_4 ~=~ \text{year 2011}$, $x_5 ~=~ \text{respect authority -1}$, $x_6 ~=~ \text{ideology}$, $x_7 ~=~ \text{God important}$, $x_8 ~=~ \text{respect authority 1}$, and $x_9 ~=~ \text{national pride}$.

@@

We will take a quick look at the 95% **confidence intervals** of the parameters from our fitted model, displayed in the following plots:

```{r ci}
#| echo: false
df <- tidy(abortion_fit, conf.int = TRUE)
intercept <- df[1,]
yr <- df[2,]
cai <- df[4,]
df <- df[-c(1,2,4),]

p1 <- ggplot(intercept, aes(x = term, y = estimate, ymin = conf.low, 
                            ymax = conf.high)) +
  geom_pointrange() +
  labs(x = NULL)
p2 <- ggplot(yr, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() +
  labs(x = NULL) +
  scale_x_discrete(labels=c("Year of Survey"))
p3 <- ggplot(cai, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange()+
  labs(x = NULL)+
  scale_x_discrete(labels=c("Child Autonomy Index"))

p4 <- ggplot(df, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange()+
  labs(x = NULL) +
  scale_x_discrete(breaks=c("godimportant","ideology","respectauthority"),
        labels=c("Importance of God", "Political Ideology", "Respect for Authority"))

title <- ggdraw() + 
  draw_label(
    "Confidence Intervals of Model Parameter Estimates",
    x = 0,
    hjust = 0
  ) +
  theme(
    plot.margin = margin(0, 0, 0, 7)
  )

top_row <- plot_grid(p1,p2,p3, ncol = 3, nrow = 1)
plot_grid(title, 
          top_row, p4, ncol = 1, rel_heights = c(0.1, 1,1))

```

We then proceed to our **hypothesis test**. As stated in our introduction, given that this issue has become highly polarized by political party, we suspect that liberal political ideology and political attitudes will correlate with belief that abortion is more justified.

@@

HA (alternative hypothesis): An individual’s political ideology (ideology), their beliefs about child autonomy (cai), the importance of God in their life (godimportant), their view on whether respect for authority is good (respectauthority), and their pride in being American (nationalpride) are significant predictors of aj, and the variables’ respective slope point estimates will not be 0.

@@

We now consider the strength of our model in **predicting** the attitude toward the justifiability of abortion of a person of certain characteristics.

Earlier, we guessed that liberal ideology and liberal-leaning attitudes on other social issues will correlate with belief that abortion is more justified. For a person surveyed in 2011 who identifies with a left leaning political ideology ("1"), has a high child autonomy index of 2, considers God and religion unimportant ("1"), and has little respect for authority, our our model predicts that their attitude toward the justifiability of abortion is 9.27, that is to say, very amenable to the justifiability of abortion. This supports our hypothesis based on contextual knowledge of this issue.

```{r predict-new}
#| echo: false
new_person <- tibble(year = 2011, ideology = 1, cai = 2, godimportant = 1, respectauthority = -1)
predict(abortion_fit, new_data = new_person)

```
However, calculating the R-squared and root-mean-squared-error (RMSE) values of our model shows that our model is not very strong at predicting attitude toward the justifiability of abortion, with a R^2 of 0.247 and RMSE of 2.53.

@@

```{r metrics}
#| echo: false
abortion_training_pred <- predict(abortion_fit, abortion_training) %>%
  bind_cols(abortion_training %>% select(aj, year, ideology, cai, 
                                         godimportant, respectauthority))
rsq(abortion_training_pred, 
                 truth = aj, 
                 estimate = .pred)
rmse(abortion_training_pred,
                 truth = aj,
                 estimate = .pred)
```

@@

```{r cv}
#| echo: false
set.seed(206)
folds <- vfold_cv(abortion_training, v = 10)
abortion_fit_rs <- abortion_wflow %>%
  fit_resamples(folds)

@@

```{r cv-table}
#| echo: false
cv_metrics <- collect_metrics(abortion_fit_rs, summarize = FALSE) 
cv_metrics2 <- collect_metrics(abortion_fit_rs2, summarize = FALSE) 
cv_metrics3 <- collect_metrics(abortion_fit_rs3, summarize = FALSE)

@@

```{r aic-bic}
#| echo: false
glance(abortion_fit) %>% 
  select(AIC, BIC)
